Optimal design for on-farm strip trials --- systematic or randomised?

### 2023-06-22 main.tex
abstract updated

\usepackage[url=false, isbn=false, eprint=false, backref=true, style= authoryear, backend=bibtex, maxcitenames=2, giveninits=true, maxbibnames=100]{biblatex}
\addbibresource{BayesOFE.bib}

\title{Optimal design for on-farm strip trials --- systematic or randomised?}
%\author{Jerome}
\author[1]{Zhanglong Cao}
\author[1]{Andrew Grose}
\author[1]{Jordan Brown}
\author[1,2]{Suman Rakshit}

\affil[1]{SAGI West, School of Molecular and Life Sciences, Curtin University, Perth, Australia}
\affil[2]{School of Electrical Engineering, Computing and Mathematical Sciences, Curtin University, Perth, Australia}

@@

\begin{abstract}
There is no doubt on the importance of randomisation in agricultural experiments by agronomists and biometricians. Even when agronomists extend the experimentation from small trials to large on-farm trials, randomised designs predominate over systematic designs. However, the situation may change depending on the objective of the on-farm experiments (OFE). If the goal of OFE is obtaining a smooth map showing the optimal level of a controllable input across a grid made by rows and columns covering the whole field, a systematic design should be preferred over a randomised design in terms of robustness and reliability. With the novel geographically weighted regression (GWR) for OFE and simulation studies, we conclude that, for large OFE strip trials, the difference between randomised designs and systematic designs are not significant if a linear model of treatments is fitted or if spatial variation is not taken into account. But for a quadratic model, systematic designs are superior to randomised designs. 
\end{abstract}

@@

The principles of randomisation were first expounded in 1925, when \textcite{Fisher1934Statistical} analysed a few systematically arranged experiments and pointed out that randomisation can provide valid tests of significance subject to appropriate restrictions, such as experimental units arranged in blocks or in the rows and columns of a Latin square \parencite{Verdooren2020History}. The most straightforward, least restrictive experimental design is the completely randomised design (CRD). More complex designs are widely used in agricultural experiments \parencite{Petersen1994Agricultural}, which include randomised complete block, split-plot, strip-plot, and latin square designs.

@@

Following these principles, randomised designs are routinely used, from small on-site trials to large on-farm experiments (OFE), in contrast to systematic designs which are rarely used. OFE enables farmers the flexibility to implement large-scale experiments in order to test management practices on their farms \parencite{Evans2020Assessment}. The aim of OFE is to enable farmers to improve their competence of uncertainties and to take into account their existing strengths of handling translational and structural uncertainty \parencite{Cook2013Onfarm}. In the situation that the goal is to compare yield responses between management classes or to select individuals with the best performance as new market varieties, a randomised design is superior to a systematic design \parencite{Pringle2004FieldScale, Selle2019Flexible}.

@@

Randomisation has been considered a crucial prerequisite for obtaining valid statistical inferences \parencite{Piepho2013Why}, however this is not true when the goal of OFE shifts from what is desired from a conventional analysis. In the application of precision agriculture (PA), a prescription map from the experimental results is required by the variable-rate applicators (VRA) prior to the start of the operation \parencite{Pringle2004FieldScale}. Therefore, in this scenario, the goal of OFE becomes obtaining such a smooth map showing the optimal level of a controllable input, such as nitrogen rates, across a grid made by rows and columns covering the whole field. For this objective, \textcite{Pringle2004FieldScale} stress that only a single level of the treatments can be directly observed at any one point on the grid, and the response for other levels at the same grid must be interpolated. If a randomised design is conducted, the interpolation distances to locations with treatment levels of interest will vary across the field. Such heterogeneous distances increase the uncertainty in the analysis and reduce the efficiency of local prediction \parencite{Piepho2011Statistical}. Hence, for this scenario a systematic design is preferable to a randomised design. Unfortunately, this standpoint has been ignored by researchers, meaning that randomised designs have been universal.

To analyse a systematic design for the creation of the optimal treatment map for OFE is statistically challenging. The truly localised estimation at each point on the grid is unknown and the optimum treatment response continuously varies spatially on the field. \textcite{Cao2021Bayesian} implemented a Bayesian approach of spatially correlated random parameters for large systematic OFE strip trials. They assumed a quadratic response in the model, consisting of a global component and a local spatially varying component. However, the Bayesian approach is time-consuming and requires preliminary knowledge of Bayesian statistics for farmers and agronomists. Alternatively, \textcite{Rakshit2020Novel} adapted a local regression approach, called geographically weighted regression (GWR), to obtain spatially-varying estimates of treatment effects for OFE. Additionally, \textcite{Evans2020Assessment} conclude that, through simulation studies, GWR is a simple method for the analysis of OFE data and is able to accurately separate yield variation that is not due to the applied treatment from yield response due to treatment. The limitation in the study is that they used a randomised design and assumed a linear response to fertiliser treatment.

@@

\textcite{Piepho2018Tutorial} demonstrate an example that a linear model is lack of fit on the sugar beets data \parencite{Petersen1994Agricultural}. \textcite{Glynn2007} show that many curves exist beyond a linear trend for nutrient-response relationships. 
%However, when modelling this process it is still common to see linear approximations made \parencite{}. 
%While segments of nutrient-response relationships may be linear, they will generally exhibit inflection points which arise from processes such as sufficiency or toxicity. 
The curve depends on the current availability of other macro and micro nutrients in the soil \parencite{Marschner2011}, meaning that a linear relationship is unlikely to be consistent across a large trial. For this reason, it is important to consider models with degrees higher than 1, with a quadratic model found to be adequate relationships (\textcite{Piepho2018Tutorial, Liben2019}).

@@

Our study uses a few simulation examples to demonstrate that randomisation is not a crucial prerequisite for large OFE strip trials, and, in the purpose of obtaining a treatment map, a systematic design is superior over a randomised design subject to appropriate restrictions. We also test the power of GWR, allowing us to know if it can successfully estimate spatially varying coefficients for both linear and quadratic responses to treatments. It shows that the optimal bandwidths found by AICc is not the best bandwidth for GWR, but a fixed bandwidth based on the experimental design is recommended.

@@

\textcite{Cao2021Bayesian} implemented the Bayesian hierarchical model with spatially correlated random parameters on large OFE strip trials. Here, for the simulation study, we use the same model to generate the simulated data.

@@

\begin{figure}[H]
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Expt/Linlines.pdf}
		\caption{Linear relationship of yield to nitrogen level.}
	\end{subfigure}
	\hspace{0.05\textwidth}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Expt/Qualines.pdf}
		\caption{Quadratic relationship of yield to nitrogen level.}
	\end{subfigure}
	\caption{Noise-free linear and quadratic response to fertiliser treatment.}\label{fig:Lines}
\end{figure}

@@

We assessed the true mean squared error (MSE) of estimated coefficients differences. This was calculated by the difference of true coefficients, $\bm{\beta} = b + \bm{u}$, and estimated spatially varying coefficients, $\bm{\hat{\beta}} = \hat{b}+\bm{\hat{u}}$, for each grid, and then squared the discrepancy and averaged across the field for comparison. The results are shown in Figures \ref{fig:LinBetaMSE} and \ref{fig:QuadBetaMSE} where ``NS'' stands for no spatial variation ($V_s=I_{n\times n}$), ``AR1'' is for $\AR(0.15)\otimes \AR(0.5)$ and ``Matern'' is the \Matern covariance with $\nu=\frac{3}{2}$. Since the MSE of $\beta_0$, $\beta_1$ and $\beta_2$ are small, we take the natural logarithm for better visualisation.

@@

The relative MSE for each bandwidth was found to change according to the spatial covariance matrix as well as the coefficient. When no spatial variation was simulated for $\beta_1$ and $\beta_2$, the bandwidth selected by AICc had the lowest MSE, followed by the bandwidths 9 and then 5. When there was spatial variation (either AR1$\otimes$AR1 or \Matern) present the estimation for $\beta_1$ and $\beta_2$ had lowest MSE when using bandwidth 9, followed by 5 and then bandwidth from AICc. The intercept ($\beta_0$) showed only significant changes in MSE amongst the bandwidths when the \Matern spatial covariance was considered. In this case, the bandwidth selected by AICc produced the smallest MSE.

@@

Furthermore, ANOVA techniques were used for the analyses of the above results. The analyses were performed for two scenarios of different responses separately. For each response, the coefficients were also accounted for in the model. The objective of the analysis was to investigate the five main factors: two types of design, three bandwidths, three covariance matrices, coefficients $\bm{\beta}$ and the correlation $\epsilon$. The significance patterns of the second order interactions was also of interest. The results are listed in Table \ref{tb:LMMoutput}.

The results are consistent with what was observed in the subsection above in that, for the linear response, the difference between randomised and systematic designs was not significant. However, for the quadratic response, the design and its interactions with bandwidth and the coefficients were significant. For both scenarios, the correlation intensity and all of its interactions were not significant (Table \ref{tb:LMMoutput}). Hence, GWR performs similarly with either low or high correlation between coefficients. Also of note was that the Bandwidth, and its second order interactions with the variables other than the correlation, was found to be significant for both response types.

@@

From the simulation study, we found that the bandwidth given by AICc skewed to 1 if spatial covariance was included in the model, for all types of the design and nature of the response. If spatial covariance was not introduced, the GWR tended to use all data in one row (Figure \ref{fig:histband}).

@@

%1. Summary of Main Findings
Agronomists and biometricians generally prefer randomised designs for OFE trials. According to the performance metrics used, our simulation study shows a systematic design performed either preferably or similarly to a randomised design for the purposes of creating a varying treatment map. The differentiating factors included primarily the response type and the spatial covariance model, while the correlation amongst the treatment coefficients was not found to be important. These are factors which can be assessed by the farmer beforehand, and this should dictate which design should be used. However, given that a systematic design is easier to implement in the field, and shows little downside for the purposes of creating a varying treatment map, we will advocate for the use of systematic designs.

@@

%2. Discussing Spatial Covariance Structures
%- Some Questions that might be interesting to answer
%- Is a \Matern class more realistic and what are its limitations? %- 
Another consideration for farmers as to which design to use is the expected spatial covariance structure in the field. When no spatial structure was simulated, the differences between the prediction from the systematic and random design were minimal. This result should be expected given if there are no spatial autocorrelations, then the individual query grids are independent observations and therefore the design is not important. However, when a first-order autoregressive structure was simulated the differences were noticeable when a quadratic response was used, showing systematic designs to be preferential. The largest difference between the two designs came when considering the \Matern spatial covariance structure, which showed a clear preference for systematic designs when a quadratic response was considered, and also a small preference for systematic designs for a linear response. Therefore, only if spatial variability was predicted to be negligible in the field would using a randomised design be reasonable given a quadratic response. This assumption of negligible spatial variability would be tough to reason with given the large fields used in on-farm experimentation (add reference), meaning that in application a systematic design should be used.

@@

%5. Recommendation of design variations
Given the scope of the paper, some designs and factors were not considered. Designs such as chequerboard or wave designs have been suggested for on-farm experiments \parencite{bramley1999designing}, however weren't able to be considered here. Topographical factors (spatial zones) were also not entertained in our study. Since GWR estimates a global template model and then adjusts it at a local scale across the study region, the variation between zones is ``flushed out'' by the spatial covariance.

>>>>

\usepackage[url=false, isbn=false, eprint=false, backref=true, style= authoryear, backend=bibtex, maxcitenames=2, giveninits=true, maxbibnames=100]{biblatex}
\DeclareNameAlias{sortname}{last-first}
\addbibresource{BayesOFE.bib}

\title{Optimal design for on-farm strip trials --- systematic or randomised?}
%\author{Jerome}
\author[1,2]{Zhanglong Cao}
\author[1,2]{Andrew Grose}
\author[1,2]{Jordan Brown}
\author[1,3]{Suman Rakshit}
\author[1,2,3]{XXXX}

\affil[1]{CBADA, School of Molecular and Life Sciences, Curtin University, Perth, Australia}
\affil[2]{Centre for Crop and Disease Management, School of Molecular and Life Sciences, Curtin University, Perth, Australia}
\affil[3]{School of Electrical Engineering, Computing and Mathematical Sciences, Curtin University, Perth, Australia}

@@

\begin{abstract}
% There is no doubt on the importance of randomisation in agricultural experiments by agronomists and biometricians. Even when agronomists extend the experimentation from small trials to large on-farm trials, randomised designs predominate over systematic designs. However, the situation may change depending on the objective of the on-farm experiments (OFE). If the goal of OFE is obtaining a smooth map showing the optimal level of a controllable input across a grid made by rows and columns covering the whole field, a systematic design should be preferred over a randomised design in terms of robustness and reliability. With the novel geographically weighted regression (GWR) for OFE and simulation studies, we conclude that, for large OFE strip trials, the difference between randomised designs and systematic designs is not significant if a linear model of treatments is fitted or if the spatial variation is not taken into account. But for a quadratic model, systematic designs are superior to randomised designs. 
Randomisation is a crucial aspect of agricultural experiments for both agronomists and biometricians. While randomised designs are predominantly used for small trials and large on-farm experiments (OFE), the choice between randomised and systematic designs may depend on the objective of the OFE. If the goal is to produce a smooth map of optimal input levels across a grid covering the entire field, a systematic design may be preferred for its robustness and reliability. Our study using geographically weighted regression (GWR) and simulation shows that for large OFE strip trials, the difference between randomised and systematic designs is insignificant when fitting a linear model or ignoring spatial variation. However, for quadratic models with the presence of spatial variation, systematic designs are superior to randomised designs.
\end{abstract}

@@

The principles of randomisation were first expounded in 1925, when \textcite{Fisher1934Statistical} analysed a few systematically arranged experiments and pointed out that randomisation can provide valid tests of significance subject to appropriate restrictions, such as experimental units arranged in blocks or in rows and columns of a Latin square \parencite{Verdooren2020History}. The most straightforward, least restrictive experimental design is the completely randomised design (CRD). More complex designs, such as randomised complete block design, split-plot design, strip-plot design and latin square design, are also widely used in agricultural experiments \parencite{Petersen1994Agricultural}, Following these principles, randomised designs are routinely used from small on-site trials to large on-farm experiments (OFE) in contrast to systematic designs, which are rarely used.

@@

OFE enables farmers the flexibility to implement large-scale experiments in order to test management practices on their farms \parencite{Evans2020Assessment}. The aim of OFE is to enable farmers to improve their competence of uncertainties and to take into account their existing strengths of handling translational and structural uncertainty \parencite{Cook2013Onfarm}. In the situation that the goal is to compare yield responses between management classes or to select individuals with the best performance as new market varieties, a randomised design is superior to a systematic design \parencite{Pringle2004FieldScale, Selle2019Flexible}.

@@

Randomisation has been considered a crucial prerequisite for obtaining valid statistical inferences \parencite{Piepho2013Why}, however, this is not true when the goal of OFE shifts from what is desired from the conventional analysis. In the application of precision agriculture (PA), a prescription map from the experimental results is required by the variable-rate applicators (VRA) prior to the start of the operation \parencite{Pringle2004FieldScale}. Therefore, in this scenario, the goal of OFE becomes obtaining a smooth map showing the optimal level of a controllable input, such as nitrogen rates, across a grid made of rows and columns covering the whole field. For this objective, \textcite{Pringle2004FieldScale} stresses that only a single level of the treatments can be directly observed at any one point on the grid, and the response for other levels at the same grid must be interpolated. If a randomised design is conducted, the interpolation distances to locations with treatment levels of interest will vary across the field. Such heterogeneous distances increase the uncertainty in the analysis and reduce the efficiency of local prediction \parencite{Piepho2011Statistical}. Hence, for this scenario, a systematic design is preferable to a randomised design. Unfortunately, this standpoint has been ignored by researchers, meaning that randomised designs have been universal.

To analyse a systematic design for the creation of the optimal treatment map for OFE is statistically challenging. The truly localised estimation at each point on the grid is unknown and the optimum treatment response continuously varies spatially on the field. \textcite{Cao2022Bayesian} implemented a Bayesian approach of spatially correlated random parameters for large systematic OFE strip trials. They assumed a quadratic response in the model, consisting of a global component and a local spatially varying component. However, the Bayesian approach is time-consuming and requires preliminary knowledge of Bayesian statistics for farmers and agronomists. Alternatively, \textcite{Rakshit2020Novel} adapted a local regression approach, called geographically weighted regression (GWR), to obtain spatially-varying estimates of treatment effects for OFE. Additionally, \textcite{Evans2020Assessment} conclude that, through simulation studies, GWR is a simple method for the analysis of OFE data and is able to accurately separate yield variation that is not due to the applied treatment from yield response due to treatment. The limitation in the study is that they used a randomised design and assumed a linear response to fertiliser treatment.

@@

\textcite{Piepho2018Tutorial} demonstrate an example that a linear model is lack of fit on the sugar beets data \parencite{Petersen1994Agricultural}. \textcite{Glynn2007} show that many curves exist beyond a linear trend for nutrient-response relationships. 
%However, when modelling this process it is still common to see linear approximations made \parencite{}. 
%While segments of nutrient-response relationships may be linear, they will generally exhibit inflection points which arise from processes such as sufficiency or toxicity. 
The curve depends on the current availability of other macro and micronutrients in the soil \parencite{Marschner2011}, meaning that a linear relationship is unlikely to be consistent across a large trial. For this reason, it is important to consider models with degrees higher than 1, with a quadratic model found to be adequate relationships (\textcite{Piepho2018Tutorial, Liben2019}).

@@

Our study uses a few simulation examples to demonstrate that randomisation is not a crucial prerequisite for large strip trials, and, in the purpose of obtaining a treatment map, a systematic design is superior over a randomised design subject to appropriate restrictions. We also test the power of GWR, allowing us to know if it can successfully estimate spatially varying treatment effects for both linear and quadratic responses to treatments. It shows that the optimal bandwidth found by AICc is not the best bandwidth for GWR, but a fixed bandwidth based on the experimental design is recommended.

@@

\textcite{Cao2022Bayesian} implemented the Bayesian hierarchical model with spatially correlated random parameters on large OFE strip trials. Here, for the simulation study, we use the same model to generate the simulated data.

@@

\begin{figure}[H]
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Expt/LinlinesV2.pdf}
		\caption{Linear relationship of yield to nitrogen level.}
	\end{subfigure}
	\hspace{0.05\textwidth}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Expt/QualinesV2.pdf}
		\caption{Quadratic relationship of yield to nitrogen level.}
	\end{subfigure}
	\caption{Noise-free linear and quadratic response to fertiliser treatment.}\label{fig:Lines}
\end{figure}

@@

We assessed the true mean squared error (MSE) of estimated coefficient differences. This was calculated by the difference of true coefficients, $\bm{\beta} = b + \bm{u}$, and estimated spatially varying coefficients, $\bm{\hat{\beta}} = \hat{b}+\bm{\hat{u}}$, for each grid, and then squared the discrepancy and averaged across the field for comparison. The results are shown in Figures \ref{fig:LinBetaMSE} and \ref{fig:QuadBetaMSE} where ``NS'' stands for no spatial variation ($V_s=I_{n\times n}$), ``AR1'' is for $\AR(0.15)\otimes \AR(0.5)$ and ``Matern'' is the \Matern covariance with $\nu=\frac{3}{2}$. Since the MSE of $\beta_0$, $\beta_1$ and $\beta_2$ are small, we take the natural logarithm for better visualisation.

@@

The relative MSE for each bandwidth was found to change according to the spatial covariance matrix as well as the coefficient. When no spatial variation was simulated for $\beta_1$ and $\beta_2$, the bandwidth selected by AICc had the lowest MSE, followed by the bandwidths 9 and then 5. When there was spatial variation (either AR1$\otimes$AR1 or \Matern) present the estimation for $\beta_1$ and $\beta_2$ had the lowest MSE when using bandwidth 9, followed by 5 and then bandwidth from AICc. The intercept ($\beta_0$) showed only significant changes in MSE amongst the bandwidths when the \Matern spatial covariance was considered. In this case, the bandwidth selected by AICc produced the smallest MSE.

@@

Furthermore, ANOVA techniques were used for the analyses of the above results. The analyses were performed for two scenarios of different responses separately. For each response, the coefficients were also accounted for in the model. The objective of the analysis was to investigate the five main factors: two types of design, three bandwidths, three covariance matrices, coefficients $\bm{\beta}$ and the correlation $\epsilon$. The significance patterns of the second-order interactions was also of interest. The results are listed in Table \ref{tb:LMMoutput}.

The results are consistent with what was observed in the subsection above in that, for the linear response, the difference between randomised and systematic designs was not significant. However, for the quadratic response, the design and its interactions with bandwidth and the coefficients were significant. For both scenarios, the correlation intensity and all of its interactions were not significant (Table \ref{tb:LMMoutput}). Hence, GWR performs similarly with either a low or high correlation between coefficients. Also of note was that the Bandwidth, and its second-order interactions with the variables other than the correlation, was found to be significant for both response types.

@@

From the simulation study, we found that the bandwidth given by AICc skewed to 1 if spatial covariance was included in the model, for all types of the design and the nature of the response. If spatial covariance was not introduced, the GWR tended to use all data in one row (Figure \ref{fig:histband}).

@@

%1. Summary of Main Findings
Agronomists and biometricians generally prefer randomised designs for OFE trials. According to the performance metrics used, our simulation study shows a systematic design performed either preferably or similarly to a randomised design for the purposes of creating a varying treatment map. The differentiating factors included primarily the response type and the spatial covariance model, while the correlation amongst the treatment coefficients was not found to be important. These are factors that can be assessed by the farmer beforehand, and this should dictate which design should be used. However, given that a systematic design is easier to implement in the field, and shows little downside for the purposes of creating a varying treatment map, we will advocate for the use of systematic designs.

@@

%2. Discussing Spatial Covariance Structures
%- Some Questions that might be interesting to answer
%- Is a \Matern class more realistic and what are its limitations? %- 
Another consideration for farmers as to which design to use is the expected spatial covariance structure in the field. When no spatial structure was simulated, the differences between the prediction from the systematic and random design were minimal. This result should be expected given if there are no spatial autocorrelations, then the individual query grids are independent observations and therefore the design is not important. However, when a first-order autoregressive structure was simulated the differences were noticeable when a quadratic response was used, showing systematic designs to be preferential. The largest difference between the two designs came when considering the \Matern spatial covariance structure, which showed a clear preference for systematic designs when a quadratic response was considered, and also a small preference for systematic designs for a linear response. Therefore, only if spatial variability was predicted to be negligible in the field would using a randomised design be reasonable given a quadratic response. This assumption of negligible spatial variability would be tough to reason with given the large fields used in on-farm experimentation (add reference), meaning that in the application a systematic design should be used.

@@

%5. Recommendation of design variations
Given the scope of the paper, some designs and factors were not considered. Designs such as chequerboard or wave designs have been suggested for on-farm experiments \parencite{bramley1999designing}, however, were not able to be considered here. Topographical factors (spatial zones) were also not entertained in our study. Since GWR estimates a global template model and then adjusts it at a local scale across the study region, the variation between zones is ``flushed out'' by the spatial covariance.
